---
title: "SpeciesPortfolio"
author: "Kiri Daust"
date: "17/05/2020"
output:
  html_document: default
  pdf_document: default
---

This script uses an adaptation of Markowitz portolio method to calculate the optimal mix of tree species, based on modelling species growth using site predictions from the CCISS tool, and estimating SIBEC data. The main optimisation functions are sourced from a C++ and Python script.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(ggplot2)
require(MASS)
require(magrittr)
require(foreach)
require(reshape2)
require(doParallel)
require(reticulate)
require(Rcpp)
library(gridExtra)
library(data.table)
library(scales)
library(tidyr)
library(magrittr)
require(zoo)
```

Source functions from C and Py
```{r read in  portfolio functions}
sourceCpp("./CppFunctions/SimGrowth.cpp")
source_python("./PythonFns/PortfolioOptimisation.py")
```

Load data
```{r load species data}
###Read in data
SuitTable <- fread("InputsGit/Feasibility_v11_21.csv") ##tree spp suitability
SuitTable <- unique(SuitTable)

colnames(SuitTable)[2:4] <- c("SS_NoSpace","Spp","Suitability")

SIBEC <- fread("InputsGit/BartPredSI.csv") ###import SI data (currently from BART)
SIBEC <- SIBEC[,-4] %>% 
  set_colnames(c("SS_NoSpace", "TreeSpp","MeanPlotSiteIndex"))

SSPredAll <- fread("InputsGit/FireSSPred.csv") ##Import SS predictions from CCISS tool: must have columns MergedBGC, Source, SS_NoSpace, SSprob, SSCurrent, FuturePeriod, SiteNo
SSPredAll <- SSPredAll[,c("MergedBGC", "Source", "SS_NoSpace", "SSprob", "SSCurrent", 
                          "FuturePeriod", "SiteNo")]
# selectBGC <- select.list(choices = sort(unique(SSPredAll$SSCurrent)), graphics = TRUE) ###Select BGC to run for
selectBGC <- "IDFdk3/01"
SSPredAll <- SSPredAll[SSCurrent == selectBGC,]

SSPredSave <- SSPredAll
#SSPredAll <- SSPredSave

SiteList <- unique(SSPredAll$SiteNo[grep("Ele",SSPredAll$SiteNo)]) ###select Elephant hill fire
SSPredAll <- SSPredAll[SiteNo %in% SiteList & !is.na(SSPredAll$SSprob),]
SList <- unique(SSPredAll$SiteNo)
```

Function to clean and summarise SIBEC and CCISS data
```{r CCISS predictions}
##function to clean data and summarise 
cleanData <- function(SNum,Trees){
  SSPred <- SSPredAll[SiteNo == SNum,] ###subset
  
  ##Merge SIBEC data
  SIBEC <- SIBEC[TreeSpp %in% Trees,]
  SSPred <- SSPred[,c(6,3,4)]
  SSPred <- merge(SSPred, SIBEC, by = "SS_NoSpace", all.x = TRUE)
  
  
  ###Add rows for species with missing SI - mostly US units here
  add <- foreach(Year = unique(SSPred$FuturePeriod), .combine = rbind) %do%{
    byYear <- SSPred[SSPred$FuturePeriod == Year,]
    foreach(SS = unique(byYear$SS_NoSpace), .combine = rbind) %do%{
      bySS <- byYear[byYear$SS_NoSpace == SS,]
      missing <- Trees[!Trees %in% bySS$TreeSpp]
      new <- bySS[rep(1,length(missing)),]
      new$TreeSpp <- missing
      new
    }
  }
  if(nrow(add) > 0){
    add$MeanPlotSiteIndex <- 5 ##Set missing SI
    SSPred <- rbind(SSPred, add)
  }
  
  
  SSPred <- SSPred[!is.na(SSPred$TreeSpp),]
  colnames(SSPred)[4] <- "Spp"
  
  ##Add suitability
  SSPred <- merge(SSPred, SuitTable, by = c("SS_NoSpace","Spp"), all.x = TRUE)
  SSPred$Suitability[is.na(SSPred$Suitability)] <- 5
  
  ###Create current data
  current <- SIBEC[SIBEC$SS_NoSpace == selectBGC,] %>% 
    merge(SuitTable, by.x = c("TreeSpp","SS_NoSpace"), by.y = c("Spp","SS_NoSpace"), all.x = TRUE) %>%
    unique()
  
  ###check that there aren't errors in the table
  temp <- aggregate(SS_NoSpace ~ TreeSpp, current, FUN = length)
  if(any(temp$SS_NoSpace > 1)){
    stop("There are partial duplicates in the suitablity table. Please fix them. :)")
  }
  
  current <- data.frame(Spp = current$TreeSpp, FuturePeriod = 2000, MeanSI = current$MeanPlotSiteIndex, MeanSuit = current$Suitability)
  
  missing <- Trees[!Trees %in% current$Spp]
  if(length(missing) > 0){
    new <- current[rep(1,length(missing)),]
    new$Spp <- missing
    new$MeanSI <- 10
    new$MeanSuit <- 5
    current <- rbind(current, new)
  }
  
  ##Summarise data- average SI and Suit weighted by SSProb
  SS.sum <- SSPred[,.(MeanSI = sum(MeanPlotSiteIndex*(SSprob/sum(SSprob))),
                      MeanSuit = round(sum(Suitability*(SSprob/sum(SSprob))), digits = 0)),
                   by = .(Spp,FuturePeriod)]
  
  SS.sum <- rbind(SS.sum, current)
  ###not sure what we were doing here?
  SS.sum$MeanSI[SS.sum$MeanSuit == 4] <- 5
  SS.sum$MeanSI[SS.sum$MeanSuit == 5] <- 0
  SS.sum$MeanSuit[SS.sum$MeanSuit == 5] <- 4
  SS.sum <- unique(SS.sum)
  SS.sum <- SS.sum[order(SS.sum$Spp,SS.sum$FuturePeriod),]
  return(SS.sum)
}

combineList <- function(...) {##Combine multiple dataframe in foreach loop
  mapply(FUN = rbind, ..., SIMPLIFY=FALSE)
}

```

Specify parameters...
```{r model parameters}
### Probabilities attached to each suitability
SuitProb <- data.frame("Suit" = c(1,2,3,4), "ProbDead" = c(0.5,0.5,1,4), 
                       "NoMort" = c(70,60,50,30)) ####ProbDead- out of 100 trees, how many will die each year at each suitability. NoMort- Percent of time no mortality

Trees <- c("At","Bl","Cw","Fd","Hw","Lw","Pl","Py","Sx") ##set species to use in portfolio
nSpp <- length(Trees)
treeList <- Trees
##set min and max weights for each species
# minWt <- c(0,0,0,0,0.05,0.2,0,0.1)
# maxWt <- c(0.3,0.3,0.8,0.5,0.4,0.6,1,1)
# boundDat <- data.frame(Spp = Trees, Min = minWt, Max = maxWt)
minWt <- 0
maxWt <- 1
minAccept <- 0.1 ##min acceptable weight in porfolio - if lower, will remove and re-optimize

```

Run optimisation (note this may take a bit)

```{r run portfolio}
SL <- SList#[1:20]
allSitesSpp <- foreach(SNum = SL, .combine = rbind, 
                       .packages = c("foreach","reshape2","dplyr","magrittr","PortfolioAnalytics", "Rcpp"), 
                       .noexport = c("simGrowthCpp")) %do% {
                         cat("Optimising site",SNum,"...\n")
                         SS.sum <- cleanData(SNum, Trees)
                         
                         cols <- rainbow(length(treeList))
                         annualDat <- data.frame("Year" = seq(2000,2100,1))
                         
                         #plot(0,0, type = "n", xlim = c(1,100), ylim = c(0,3000), xlab = "Year", ylab = "Volume")###plot
                         
                         eff_front <- foreach(w = 1, .combine = combineList) %do% { ##number of iterations
                           output <- data.frame("year" = annualDat$Year)
                           # growthSim <- data.frame(Spp = character(), Year = numeric(), Returns = numeric())
                           
                           for (k in 1:nSpp){ ##for each tree
                             
                             DatSpp <- SS.sum[SS.sum$Spp == treeList[k],]
                             
                             dat <- data.frame("Period" = c(2000,2025,2055,2085), "SIBEC" = DatSpp$MeanSI, "Suit" = DatSpp$MeanSuit)
                             dat$SIBEC <- dat$SIBEC/50 ##for mean annual increment
                             dat <- merge(dat, SuitProb, by = "Suit")
                             s <- approx(dat$Period, dat$SIBEC, n = 101) ##Smooth SI
                             p <- approx(dat$Period, dat$ProbDead, n = 101) ###Smooth Prob Dead
                             m <- approx(dat$Period, dat$NoMort, n = 101) ##Smooth No Mort
                             
                             ###data frame of annual data
                             annualDat <- data.frame("Year" = seq(2000,2100,1), "Growth" = s[["y"]], "MeanDead" = p[["y"]], "NoMort" = m[["y"]]) ##create working data
                             
                             Returns <- simGrowthCpp(DF = annualDat)
                             #lines(Returns, type = 'l', col = cols[k])##plot
                             tmpR <- c(0,Returns)
                             assets <- Returns - tmpR[-length(tmpR)]
                             temp <- data.frame(Spp = rep(treeList[k],101), Year = 1:101, Returns = Returns)
                             output <- cbind(output, assets)
                           } ## for each tree species
                           
                           #legend("topleft", treeList, col = cols, pch = 15)
                           
                           colnames(output) <- c("Year", treeList)
                           
                           ####Portfolio#######################################
                           returns <- output
                           rownames(returns) <- returns[,1]
                           returns <- returns[,-1]
                           ###only include species with mean return > 1 in portfolio
                           use <- colnames(returns)[colMeans(returns) > 1] ###should probably be higher
                           returns <- returns[,use]
                           sigma2 <- as.data.frame(cor(returns)) ###to create cov mat from returns
                       
                           ef <- ef_weights_v2(returns, sigma2, minWt,maxWt,minAccept) 
                           ef_w <- ef[[1]]
                           ef_w$Sd <- ef[[2]]
                           ef_w$Return <- 1:20
                           ef_w$It <- w
                           
                    
                           list(frontier = ef_w)
                         }
                         
                         eff_front2 <- as.data.table(eff_front[['frontier']])
                         eff_front2$SiteNo <- SNum
                         melt(eff_front2, id.vars = c("Return","It","SiteNo"),variable.name = "Spp")
                       }
```

Plot by volatility...

```{r plot portfolio curves}
require(ggthemes)

efAll <- allSitesSpp
efAll <- efAll[,-c("It")]
efAll <- dcast(efAll,Return ~ Spp, fun.aggregate = function(x){sum(x)/(length(SL))})
efAll <- efAll[complete.cases(efAll),]
efAll$Return <- efAll$Return/max(efAll$Return) ##standardise return

efAll <- melt(efAll, id.vars = "Sd")

myColours <- c("brown","red","pink", "orange","yellow","green","blue","magenta","darkgoldenrod")
names(myColours) <- levels(factor(Trees))
colScale <- scale_fill_manual(name = "variable", values = myColours)

ggplot(efAll[efAll$variable != "Return",])+
  geom_area(aes(x = Sd, y = value, fill = variable),size = 0.00001, col = "black", stat = "identity")+
  colScale +
  geom_line(data = efAll[efAll$variable == "Return",], aes(x = Sd, y = value))+
  scale_x_reverse() +
  xlab("Volatility")+
  ggtitle(selectBGC)+
  theme_few()
  #scale_colour_few()

```
