---
title: "Climate Change Optimized Species Portfolio for Reforestation Planning:Sunshine Coast"
author: "Will MacKenzie and Kiri Daust"
date: "17/05/2020"
output:
  #bookdown::html_document2:
    html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
  
---

This report presents a range of risk-return optimized portfolios of tree species generated using Markowitz Modern Portfolio Theory (MPT). The analysis uses forecast site series level changes to species feasibility from the Climate Change-Informed Species Selection (CCISS) tool, and site index by site series growth information (SIBEC) to calculate the optimal mix of tree species given the modelled uncertainty in future climates up to 2070.
The CCISS model forecasts future BGCs for 30 GCM/carbon scenarios for all locations, and aligns the current site series with future potential site series via edatopic overlap. We use a monte carlo approach to simulate growth of each tree species using site index data and stochastic losses based on predicted suitability. The portfolio analysis uses the modelled growth rate and a species covariance matrix to optimise the efficient frontier.
User inputs to the script are spatial point locations with climate data that will represent BGC-based analysis units (e.g. MHmm1_high and MHmm1_low). Users can also adjust the min and max values for each species and the minimum value to be included in the portfolio. 
This report is generated from Rmarkdown, and sources functions from python and C++.

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
require(dplyr)
require(ggplot2)
require(MASS)
require(magrittr)
require(foreach)
require(reshape2)
require(reticulate)
require(Rcpp)
library(gridExtra)
library(data.table)
library(scales)
library(tidyr)
library(magrittr)
library(ggthemes)
library(flextable)
require(rgdal)
require(tmap)
library(sf)
require(tufte)
require(Hmisc)
require(janitor)
require(tidyverse)
require(Rmisc)
require(bookdown)
```



```{r read in  portfolio functions}
##Source functions for growth simulations (C++), Portfolio optimization (Python) and CCISS analysis (R)
sourceCpp("./CppFunctions/SimGrowth.cpp")
source_python("./PythonFns/PortfolioOptimisation.py")
source("CCISS_Fns.R")
```



```{r load species data}
##Load data and run CCISS functions (if necessary)

inputDatName <- "SunshPntsNew_90 GCMsMSY.csv" ##generated in ClimateBC for all variable and all future climate scenarios

##Set drive with cloud data
if(dir.exists("C:/users/whmacken/Sync")){
  cloud_dir <- "C:/users/whmacken/Sync/CCISS_data/"
}else{
  cloud_dir <- "C:/Users/kirid/Sync/CCISS_data/"
}

###Read in data
SuitTable <- fread("InputsGit/Feasibility_v11_21.csv") ##tree spp suitability
SuitTable <- unique(SuitTable)
SuitNew <- fread("InputsGit/Feas_toAdd.csv")

colnames(SuitTable)[2:4] <- c("SS_NoSpace","Spp","Suitability")
SuitTable <- SuitTable[,c("BGC","SS_NoSpace","Spp","Suitability")]
SuitTable <- rbind(SuitTable, SuitNew)

SIBEC <- fread("InputsGit/PredSI_May2020.csv") 
SIBECnew <- fread("InputsGit/SI_to_add.csv")
SIBEC <- rbind(SIBEC, SIBECnew)
###import SI data (currently from BART)

SIBEC <- SIBEC[,c("SS_NoSpace","Spp","SIPred")] %>% set_colnames(c("SS_NoSpace","TreeSpp","MeanPlotSiteIndex"))

eda <- fread("InputsGit/Edatopic_v11_20.csv")
eda <- eda[,.(BGC,SS_NoSpace, Edatopic)]

if(file.exists(paste0(cloud_dir,"CCISSPred.Rdata"))){
  load(paste0(cloud_dir,"CCISSPred.Rdata"))
}else{
  ### run CCISS function to predict Site Series
  load(paste0(cloud_dir, "WNAv11_35_VAR_SubZone_ranger.Rdata"))
  Edatope <- fread("./InputsGit/Edatopic_v11_20.csv",data.table = F)
  rawDat <- fread(paste0(cloud_dir,"PortfolioFiles/",inputDatName),data.table = F)
  CCISSPred <- CCISS_Spp(Y1 = rawDat,BGCmodel = BGCmodel,E1 = Edatope)
  save(CCISSPred, file = paste0(cloud_dir,"CCISSPred.Rdata"))
}

SSPredOrig <- as.data.table(CCISSPred[[1]])
colnames(SSPredOrig)[1] <- "MergedBGC"
analUnits <- fread(paste0(cloud_dir,"PortfolioFiles/AnalysisUnits.csv"))

```

### Map of input locations

```{r location map setup, message = FALSE, echo= FALSE, include=FALSE}

## Create map of test locations by analysis unit
locations <- fread("./InputsGit/SunshineCoast_SamplePoints_New.csv") 
locations2 <- locations %>% dplyr::rename(latitude = "ycoord", longitude = "xcoord", elevation = "elevation_1") %>%
  dplyr::select (id, BGC_analysis,latitude,longitude)
BGCmap <- st_read(dsn = paste0(cloud_dir, "PortfolioFiles/SunshineCoast.gpkg"), layer = "BGC_Dissolved")

locations2 <- st_as_sf(locations2, coords = c("longitude","latitude"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(BGCmap))

```

```{r location map}
#tmaptools::palette_explorer() ###tool for choosing colours

tmap_mode("view")
tmap_options(basemaps = "OpenTopoMap")
tBGC <- tm_shape (BGCmap) + tm_polygons("BGC", alpha = .9)+
    tm_shape(locations2) + tm_symbols(size = 0.1, shapes = 21:25, col = "BGC_analysis")+
  tm_layout(legend.outside = TRUE, legend.outside.position = "right", legend.stack = "vertical", title = "Sample locations for portfolio development in the Sunshine Coast District", main.title.size = 1)
  
tBGC

```
  

```{r colours}
##Colour development for species 

cols <- fread("./InputsGit/PortfolioSppColours.csv")
cols <- cols[HexColour != "",]
myPal <- cols$HexColour
names(myPal) <- cols$TreeCode


library(unikn)
#seecol(myPal, hex = T, rgb = F)
```



```{r CCISS predictions}
#Function to clean and summarise SIBEC and CCISS data into MeanSI and MeanSuitability
##function to clean data and summarise 
cleanData <- function(SSPredAll,SIBEC,SuitTable,SNum,Trees,timePer){
  SSPred <- SSPredAll[SiteNo == SNum,] ###subset
  
  ##Merge SIBEC data
  SIBEC <- SIBEC[TreeSpp %in% Trees,]
  SSPred <- SSPred[SSPred$FuturePeriod %in% timePer,]
  SSPred <- SSPred[,c(6,3,4)]
  SSPred <- merge(SSPred, SIBEC, by = "SS_NoSpace", all.x = TRUE)
  
  
  ###Add rows for species with missing SI - mostly US units here
  add <- foreach(Year = unique(SSPred$FuturePeriod), .combine = rbind) %do%{
    byYear <- SSPred[SSPred$FuturePeriod == Year,]
    foreach(SS = unique(byYear$SS_NoSpace), .combine = rbind) %do%{
      bySS <- byYear[byYear$SS_NoSpace == SS,]
      missing <- Trees[!Trees %in% bySS$TreeSpp]
      new <- bySS[rep(1,length(missing)),]
      new$TreeSpp <- missing
      new
    }
  }
  if(!is.null(add)){
    if(nrow(add) > 0){
      add$MeanPlotSiteIndex <- 5 ##Set missing SI
      SSPred <- rbind(SSPred, add)
    }
  } 
  
  
  SSPred <- SSPred[!is.na(SSPred$TreeSpp),]
  colnames(SSPred)[4] <- "Spp"
  
  ##Add suitability
  SSPred <- merge(SSPred, SuitTable, by = c("SS_NoSpace","Spp"), all.x = TRUE)
  SSPred$Suitability[is.na(SSPred$Suitability)] <- 5
  
  temp <- SIBEC[SIBEC$SS_NoSpace == selectBGC,]
  if(nrow(temp) == 0){
    return(NULL)
  }
  ###Create current data
  current <- temp %>% 
    merge(SuitTable, by.x = c("TreeSpp","SS_NoSpace"), by.y = c("Spp","SS_NoSpace"), all.x = TRUE) %>%
    unique()
  
  ###check that there aren't errors in the table
  temp <- aggregate(SS_NoSpace ~ TreeSpp, current, FUN = length)
  if(any(temp$SS_NoSpace > 1)){
    stop("There are partial duplicates in the suitablity table. Please fix them. :)")
  }
  
  current <- data.frame(Spp = current$TreeSpp, FuturePeriod = 2000, 
                        MeanSI = current$MeanPlotSiteIndex, MeanSuit = current$Suitability)
  
  missing <- Trees[!Trees %in% current$Spp]
  if(length(missing) > 0){
    new <- current[rep(1,length(missing)),]
    new$Spp <- missing
    new$MeanSI <- 10
    new$MeanSuit <- 5
    current <- rbind(current, new)
  }
  
  ##Summarise data- average SI and Suit weighted by SSProb
  SS.sum <- SSPred[,.(MeanSI = sum(MeanPlotSiteIndex*(SSprob/sum(SSprob))),
                      MeanSuit = round(sum(Suitability*(SSprob/sum(SSprob))), digits = 0)),
                   by = .(Spp,FuturePeriod)]
  
  SS.sum <- rbind(SS.sum, current)
  SS.sum <- SS.sum[SS.sum$FuturePeriod %in% timePer,]
  ###not sure what we were doing here?
  SS.sum$MeanSI[SS.sum$MeanSuit == 4] <- 5
  SS.sum$MeanSI[SS.sum$MeanSuit == 5] <- 0
  SS.sum$MeanSuit[SS.sum$MeanSuit == 5] <- 4
  SS.sum <- unique(SS.sum)
  SS.sum <- SS.sum[order(SS.sum$Spp,SS.sum$FuturePeriod),]
  #SS.sum$MeanSuit[is.na(SS.sum$MeanSuit)] <- 3
  return(SS.sum)
}

combineList <- function(...) {##Combine multiple dataframe in foreach loop
  mapply(FUN = rbind, ..., SIMPLIFY=FALSE)
}


edatopicSubset <- function(SSPredOrig, analUnits, eda, pos = "Zonal"){
  if(pos == "Zonal"){
    SSPredFull <- SSPredOrig[grep("01",SSPredOrig$SSCurrent),]
  }else{
    edaSub <- eda[Edatopic == pos,]
    SSPredFull <- SSPredOrig[SSCurrent %in% edaSub$SS_NoSpace,]
  }
  SSPredFull$CurrBGC <- gsub("/.*","", SSPredFull$SSCurrent)
  
  SSPredFull <- SSPredFull[,c("MergedBGC", "Source", "SS_NoSpace", "SSprob", "SSCurrent", 
                            "FuturePeriod", "SiteNo","CurrBGC")]
  
  ##remove cases where not all timeperiods available
  SSPredFull <- as.data.table(SSPredFull)
  temp <- SSPredFull[,.(Num = length(unique(FuturePeriod))), by = c("SiteNo","SSCurrent")]
  temp <- temp[Num == 3,-c("Num")]
  SSPredFull <- SSPredFull[temp,on = c("SiteNo","SSCurrent")]
  SSPredFull[,SiteNo := as.numeric(SiteNo)]
  
  
  setkey(SSPredFull, SiteNo)
  setkey(analUnits, ID1)
  SSPredFull <- SSPredFull[analUnits]
}

loopCombine <- function(a,b){
  G <- rbind(a$GraphDat,b$GraphDat)
  df <- rbind(a$MaxS,b$MaxS)
  return(list(GraphDat = G, MaxS = df))
}

ef_plot <- function(outAll){
  efAll <- outAll$GraphDat
  intDat <- outAll$MaxS
  ggplot(efAll[efAll$variable != "RealRet",],aes(x = Sd, y = value,group = variable))+
    geom_area(aes(fill = variable), size = 0.00001, col = "grey50", stat = "identity")+
    colScale +
    geom_vline(data = intDat[intDat$Spp == "Sd",], aes(xintercept = value,colour = "blue"), 
               linetype = "twodash", size = .75)+
    geom_vline(data = intDat[intDat$Spp == "Sd",], aes(xintercept = SetRet,colour = "grey52"),
               linetype = "dashed", size = .75)+
    geom_line(data = efAll[efAll$variable == "RealRet",], 
              aes(x = Sd, y = value,colour = "black"),linetype = "F1",size = .75)+
    scale_colour_identity(name = "", guide = 'legend', labels = c("Return","MaxSharpe","90%"))+
    scale_x_reverse() +
    xlab("Max Return --> Minimized Risk")+
    ylab("Portfolio Ratio")+
    guides(fill=guide_legend("Species"))+
    theme_few()+
    facet_wrap(.~Unit, scales = "free_x")
}

MS_plot <- function(Current,Future){
  curr <- as.data.table(Current$MaxS)
  fut <- as.data.table(Future$MaxS)
  colnames(curr)[1] <- "CurrentMS"
  colnames(fut)[c(1,5)] <- c("FutureMS","Future90%")
  
  dat <- curr[fut, on = c("Unit","Spp")]
  dat <- dat[,.(Unit,Spp,CurrentMS,FutureMS,`Future90%`)]
  dat <- melt(dat, id.vars = c("Unit","Spp"))
  dat <- dat[!Spp %in% c("Sd","RealRet"),]
  
  ggplot(dat, aes(x = variable, y = value, group = Spp, fill = Spp))+
    geom_bar(stat = "identity")+
    colScale+
    facet_wrap(.~Unit)+
    labs(x = "Portfolio Type", y = "Portfolio Ratio")+
    guides(fill=guide_legend("Species"))+
    theme_few()+
    theme(axis.text.x = element_text(angle = 90))
}

```



```{r model parameters, out.width=8, results = 'hold'}
### Probabilities attached to each suitability
SuitProb <- data.frame("Suit" = c(1,2,3,4), "ProbDead" = c(0.5,0.5,1,4), 
                       "NoMort" = c(70,60,50,30)) ####ProbDead- out of 100 trees, how many will die each year at each suitability. NoMort- Percent of time no mortality

# ft1 <- flextable(SuitProb) %>% set_caption("Probability parameters for tree death")
# 

##Set minimum weighting for any species to appear in portfolio
minAccept <- 0.01 ##min acceptable weight in porfolio - if lower, will remove and re-optimize

# cat (paste (minAccept, "is the minimum percent of portolio for a species to be used in final portfolio"))

Trees <- unique(SIBEC$TreeSpp)
myColours <- data.table(TreeCode = Trees)
myColours <- cols[myColours, on = "TreeCode"]
myColours <- myColours[!is.na(HexColour),]
pal <- myColours$HexColour
names(pal) <- myColours$TreeCode
colScale <- scale_fill_manual(name = "variable", values = pal)
Trees <- myColours$TreeCode

##set min and max weights for each species
boundDat <- data.table(Spp = Trees)
boundDat[,`:=`(minWt = 0, maxWt = 1)]

##adjust min and max weights (or remove species)
## see below examples
# boundDat[Spp == "Hw",maxWt := 0] ##remove Hw
# boundDat[Spp == "Pl",minWt := 0.1] ## set min weight for Pl to 0.1

boundDat2 <-  data.table::transpose(boundDat) %>% row_to_names (row_number = 1)

##############
ft2 <- flextable(boundDat2) %>% set_caption("Minimum and max percent of portolio weighting for individual species in portfolio. This values can be set by the user and the portfolio will optimize around the limiate")

notUse <- boundDat$Spp[boundDat$maxWt == 0]
if(length(notUse > 0)){
  Trees <- Trees[!Trees %in% notUse]
}
nSpp <- length(Trees)
treeList <- Trees
```




```{r parallelSetup, results = 'hide'}
##Setup for parallel computing

worker.init <- function(){
    Rcpp::sourceCpp("./CppFunctions/SimGrowth.cpp")
}

require(doParallel)
cl <- makePSOCKcluster(detectCores()-2)
clusterCall(cl, worker.init)
registerDoParallel(cl)
```


```{r portfolio_function, message = TRUE}

run_portfolio <- function(SSPredFull,SIBEC,SuitTable,SuitProb,Trees,boundDat,minAccept,timePeriods,returnValue){
  treeList <- Trees
  nSpp <- length(Trees)
  Units <- unique(SSPredFull$BGC_analysis)
  outAll <- foreach(BGC = Units, .combine = loopCombine,
                    .packages = c("reshape2","Rcpp","magrittr","reticulate",
                                  "data.table","foreach","ggplot2","ggthemes","scales"), 
                    .noexport = c("gs2gw", "simGrowthCBST","simGrowthCpp"),
                    .export = c("cleanData","loopCombine")) %dopar% {
    reticulate::source_python("./PythonFns/PortfolioOptimisation.py") ##comment out this line if running sequentially
    SSPredBGC <- SSPredFull[BGC_analysis == BGC,-("BGC_analysis")]
    SSList <- unique(SSPredBGC$SSCurrent)
    SSout <- foreach(selectBGC = SSList[1], .combine = loopCombine) %do% {
      cat("########## Processing", selectBGC,"###################\n")
      SSPredAll <- SSPredBGC[SSPredBGC$SSCurrent == selectBGC,]
      
      SiteList <- unique(SSPredAll$SiteNo)
      #SiteList <- rep(SiteList, each = round(15/length(SiteList)))
      SSPredAll <- SSPredAll[SSPredAll$SiteNo %in% SiteList & !is.na(SSPredAll$SSprob),]
      
      SL <- SiteList
      allSitesSpp <- foreach(SNum = SL, .combine = rbind, 
                         .packages = c("foreach","reshape2","dplyr","magrittr","Rcpp"), 
                         .noexport = c("simGrowthCpp")) %do% {
                           
                           #cat("Optimising site",SNum,"...\n")
                           SS.sum <- cleanData(SSPredAll,SIBEC,SuitTable,SNum, Trees, timePer = timePeriods)
                           SS.sum$FuturePeriod <- as.numeric(SS.sum$FuturePeriod)
                           if(length(timePeriods) == 1){
                             temp <- SS.sum
                             temp$FuturePeriod <- SS.sum$FuturePeriod[1]+85
                             SS.sum <- rbind(SS.sum, temp)
                           }
                           
                           if(!is.null(SS.sum)){
                             annualDat <- data.frame("Year" = seq(2000,2100,1))
                           
                             output <- data.frame("year" = annualDat$Year)
                             
                             for (k in 1:nSpp){ ##for each tree
                               
                               DatSpp <- SS.sum[SS.sum$Spp == treeList[k],]
                               
                               dat <- data.frame("Period" = rescale(as.numeric(DatSpp$FuturePeriod), 
                                                                    to = c(2000,2085)), 
                                                 "SIBEC" = DatSpp$MeanSI, "Suit" = DatSpp$MeanSuit)
                               dat$SIBEC <- dat$SIBEC/50 ##for mean annual increment
                               dat <- merge(dat, SuitProb, by = "Suit")
                               s <- approx(dat$Period, dat$SIBEC, n = 101) ##Smooth SI
                               p <- approx(dat$Period, dat$ProbDead, n = 101) ###Smooth Prob Dead
                               m <- approx(dat$Period, dat$NoMort, n = 101) ##Smooth No Mort
                               
                               ###data frame of annual data
                               annualDat <- data.frame("Year" = seq(2000,2100,1), "Growth" = s[["y"]], 
                                                       "MeanDead" = p[["y"]], "NoMort" = m[["y"]]) ##create working data
                               
                               Returns <- simGrowthCpp(DF = annualDat)
                               tmpR <- c(0,Returns)
                               assets <- Returns - tmpR[-length(tmpR)]
                               temp <- data.frame(Spp = rep(treeList[k],101), 
                                                  Year = 1:101, Returns = Returns)
                               output <- cbind(output, assets)
                             } ## for each tree species
                             
                             colnames(output) <- c("Year", treeList)
                             
                             ####Portfolio#######################################
                             returns <- output
                             rownames(returns) <- returns[,1]
                             returns <- returns[,-1]
                             ###only include species with mean return > 1 in portfolio
                             use <- colnames(returns)[colMeans(returns) > 1] ###should probably be higher
                             returns <- returns[,use]
                             sigma2 <- as.data.frame(cor(returns)) ###to create cov mat from returns
                         
                             ef <- ef_weights_v2(returns, sigma2, boundDat,minAccept) 
                             ef_w <- ef[[1]]
                             ef_w$Sd <- ef[[2]]
                             ef_w$Return <- 1:20
                             ef_w$RealRet <- ef[[3]]
                             ef_w$Sharpe <- ef[[4]]
    
                             eff_front2 <- as.data.table(ef_w)
                             eff_front2$SiteNo <- SNum
                             melt(eff_front2, id.vars = c("Return","SiteNo"),variable.name = "Spp")
                           }else{
                             NULL
                           }
                           
                         }
      
      if(!is.null(allSitesSpp)){
        efAll <- allSitesSpp
        efAll <- dcast(efAll,Return ~ Spp, fun.aggregate = function(x){sum(x)/(length(SL))})
        efAll <- efAll[complete.cases(efAll),]
        efAll$RealRet <- efAll$RealRet/max(efAll$RealRet) ##standardise return
        RetCurve <- approx(efAll$RealRet,efAll$Sd,xout = returnValue)
        ret90 <- RetCurve$y
        maxSharpe <- efAll[efAll$Sharpe == max(efAll$Sharpe),-c("Return","Sharpe")]
        maxSPos <- maxSharpe$Sd
        maxSharpe <- t(maxSharpe) %>% as.data.frame() %>% 
          mutate(Spp = rownames(.)) %>% set_colnames(c("value","Spp"))
        ret90Props <- efAll[which.min(abs(efAll$RealRet - returnValue)),-c("Return","Sharpe")]
        ret90Props <- t(ret90Props) %>% as.data.frame() %>% 
          mutate(Spp = rownames(.)) %>% set_colnames(c("value","Spp"))
        maxSharpe$SSCurrent <- selectBGC
        maxSharpe$Unit <- BGC
        maxSharpe$SetRet <- ret90Props$value
        maxSharpe$SetRet[maxSharpe$Spp == "Sd"] <- ret90
        efAll <- efAll[,-c("Return","Sharpe")]
        efAll <- melt(efAll, id.vars = "Sd")
        efAll$variable <- factor(efAll$variable, levels = sort(unique(as.character(efAll$variable))))
        efAll$Unit <- BGC
      
        list(GraphDat = efAll, MaxS = maxSharpe)
      }
      
    }
                    }
  return(outAll)

}



```

```{r zonal_port}
timePeriods = c(2000,2025,2055)
returnValue = 0.9
SSPredFull <- edatopicSubset(SSPredOrig,analUnits,eda,pos = "Zonal")
outAll <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept, timePeriods,returnValue)
```

# Zonal Sites

### Climate change species portfolios by analysis unit.
Portfolio plots for neutral zonal site series in each analysis unit based on a rotation period to 2070 showing the efficient frontier species ratios across risk/return choices.
The graphs show the ratio of species for a range of return/risk compromises.

+ The left side of the graphs is the portfolio with maximum return but not accounting for any risk.
+ The right-hand side of the graph is the “risk-free” portfolio.
+ The 90% vertical line shows the portfolio position to select using an arbitrary 90% of maximum average return (site index).
+ The MaxSharpe line shows a portfolio that shows an optimized portfolio that balances risk/return based on levels of volatility, the amount of uncertainty in climate futures. 
Other "efficient" portfolios can be selected from positions along the volatility gradient based on the practioner tolerance for risk. 

```{r plot_pf, fig.height = 13, fig.width = 12, fig.cap = "Figure 2. Portfolio graphs for neutral aspect zonal site seris by analysis unit showing species ratios from the MPT efficient frontier for a rotation period up to 2070."}
# pdfOutput <- F
# if(pdfOutput){
#   pdf("PortfolioGraphs.pdf", width = 12, height = 15)
#   grid.arrange(grobs = outAll$Graph, ncol = 3)
#   dev.off()
# }

ef_plot(outAll)

```


### Portfolio of species at the max Sharpe ratio for each analysis unit.
+ RealRet is the estimated mean site index return relative to the maximum return portfolio.
+ Sd is the volatility of the climate futures with lower numbers meaning higher volatility/uncertainty


```{r table}
MSWeights <- dcast(outAll$MaxS,Unit ~ Spp)
MSWeights[MSWeights < 0.01] <- NA ## minimum percent for species to appear in portfolio
MSWeights <- MSWeights[,c(T,colSums(MSWeights[,-1], na.rm = T) > 0)]
MSWeights[is.na(MSWeights)] <- 0
cols <- c(names(MSWeights)[!names(MSWeights) %in% Trees],names(MSWeights)[names(MSWeights) %in% Trees])
tab <- flextable(as.data.frame(MSWeights), col_keys = cols)
tab <- colformat_num(tab,j = cols[-1],digits = 2)
tab
```

```{r currPort}
timePeriodsCurr = 2000
returnValue = 0.9
SSPredFull <- edatopicSubset(SSPredOrig,analUnits,eda,pos = "Zonal")
outAllCurr <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept, timePeriodsCurr,returnValue)

```

### Comparison of max sharpe ratio portfolios and 90% return portfolios for current and future climate

```{r stacked_bar, fig.width = 8.5, fig.height = 9, fig.cap = "Figure 3. Portfolio weighting for current climate (Max Sharpe) and future climate (Max Sharpe and 90% return)"}
MS_plot(outAllCurr, outAll)
```

# Dry sites (B2)

```{r portfolio_dry}
timePeriods = c(2000,2025,2055)
returnValue = 0.9
SSPredFull <- edatopicSubset(SSPredOrig,analUnits,eda,pos = "B2")
outAlldry <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept, timePeriods,returnValue)
outAlldryCurr <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept, 
                               timePeriods = 2000,returnValue)
```

```{r plot_dry, fig.height = 13, fig.width = 12}
ef_plot(outAlldry)
MS_plot(outAlldryCurr,outAlldry)
```

# Wet sites (C6)

```{r portfolio_wet}
timePeriods = c(2000,2025,2055)
returnValue = 0.9
SSPredFull <- edatopicSubset(SSPredOrig,analUnits,eda,pos = "D6")
outAllwet <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept = 0.05, timePeriods,returnValue)
outAllwetCurr <- run_portfolio(SSPredFull,SIBEC,SuitTable, SuitProb,Trees,boundDat,minAccept, 
                               timePeriods = 2000,returnValue)
```

```{r plot_wet, fig.height = 13, fig.width = 12}
ef_plot(outAllwet)
MS_plot(outAllwetCurr,outAllwet)
```

### Predicted BGC subzone/variants for each future time period from 30 climate models

Ratio of predicted BGC units for each BGC analysis unit (multiple locations) from CCISS function.

```{r BGC_plot, fig.width=10, fig.height=6, fig.cap = "Figure 4. Ratio of predicted BGCs in each analysis unit 2040-2070"}
BGCPred <- as.data.table(CCISSPred[[2]])
setkey(BGCPred, SiteNo)
BGCPred <- BGCPred[analUnits]
BGCPred <- BGCPred[,.(Num = sum(Pred.len)), keyby = .(FuturePeriod,BGC_analysis,BGC.pred)]
totNum <- BGCPred[,.(Total = sum(Num)), keyby = .(FuturePeriod,BGC_analysis)]
BGCPred <- totNum[BGCPred]
BGCPred[,Prop := Num/Total]

period <- 2055
BGCPred <- BGCPred[FuturePeriod == period & Prop > 0.05,]
BGCPred[,Prop := Num/sum(Num), by = BGC_analysis]

ggplot(BGCPred, aes(x = BGC_analysis,y = Prop, fill = BGC.pred, label = BGC.pred))+
  geom_bar(stat = 'identity')+
  geom_text(size = 3, position = position_stack(vjust = 0.5), angle = 45)+
  theme_few()+
  theme(axis.text.x = element_text(angle = 90), legend.position = "none")
```

### Estimated site index for each portfolio species of analysis units in each future period

```{r SI_check, fig.width = 10, fig.height = 24, fig.cap = "Figure 5. Mean and 95% CI for SI within Spp and Period, weighted by SSprob"}

weighted_ci <- function(x, weights, conf.level = 0.95) { ## main idea from Hack-R on Stack Overflow
    nx <- length(x)
    df <- nx - 1
    vx <- wtd.var(x, weights, normwt = F)
    mx <- weighted.mean(x, weights)
    stderr <- sqrt(vx/nx)
    tstat <- mx/stderr
    alpha <- 1 - conf.level
    cint <- qt(1 - alpha/2, df)
    cint <- tstat + c(-cint, cint)
    cint * stderr
}

gr <- foreach(unit = unique(SSPredFull$BGC_analysis),.combine = rbind) %do% {
  sub <- SSPredFull[BGC_analysis == unit,]
  tNum <- nrow(sub)
  ind <- rep(1:nrow(sub),length(Trees))
  sub <- sub[ind,]
  sub$TreeSpp <- rep(Trees, each = tNum)
  sub <- merge(sub,SIBEC,by = c("SS_NoSpace","TreeSpp"), all.x = T)
  sub[is.na(MeanPlotSiteIndex), MeanPlotSiteIndex := 0]
  sub[,SI := weighted.mean(MeanPlotSiteIndex,SSprob), by = .(FuturePeriod, TreeSpp)]
  sub[,CI_min := weighted_ci(MeanPlotSiteIndex,SSprob)[1], by = .(FuturePeriod, TreeSpp)]
  sub[,CI_max := weighted_ci(MeanPlotSiteIndex,SSprob)[2], by = .(FuturePeriod, TreeSpp)]
  sub <- sub[!is.na(TreeSpp) & !is.nan(CI_min),]
  # sub[,SI := weighted.mean(MeanPlotSiteIndex,SSprob), by = .(FuturePeriod, TreeSpp)]
  # sub[,CI_min := weighted_ci(MeanPlotSiteIndex,SSprob)[1], by = .(FuturePeriod, TreeSpp)]
  # sub[,CI_max := weighted_ci(MeanPlotSiteIndex,SSprob)[2], by = .(FuturePeriod, TreeSpp)]
  sub
}

ggplot(gr, aes(x = TreeSpp, y = SI, fill = FuturePeriod))+
    geom_bar(stat = "identity", position = "dodge")+
    geom_errorbar(aes(ymin = CI_min, ymax = CI_max), width = 0.2, position = position_dodge(width = 0.9))+
    theme_few()+
    theme(axis.line=element_line())+
    geom_hline(yintercept = c(10,20, 30), colour = "black")+
    facet_wrap(.~BGC_analysis, ncol = 1, scales = "free_x")
```

### Estimated feasibility for each portfolio species and analysis units in each future period

```{r suit_check, fig.width = 10, fig.height = 24, fig.cap = "Figure 5. Mean and 95% CI for suitability within Spp and Period, weighted by SSprob"}

gr <- foreach(unit = unique(SSPredFull$BGC_analysis),.combine = rbind) %do% {
  sub <- SSPredFull[BGC_analysis == unit,]
  tNum <- nrow(sub)
  ind <- rep(1:nrow(sub),length(Trees))
  sub <- sub[ind,]
  sub$Spp <- rep(Trees, each = tNum)
  sub <- merge(sub,SuitTable,by = c("SS_NoSpace","Spp"), all.x = T)
  sub[is.na(Suitability), Suitability := 5]
  sub[,Avg := weighted.mean(Suitability,SSprob), by = .(FuturePeriod, Spp)]
  sub <- sub[!is.na(Spp),]
  sub
}

ggplot(gr, aes(x = Spp, y = Avg, fill = FuturePeriod))+
    geom_bar(stat = "identity", position = "dodge")+
    theme_few()+
    theme(axis.line=element_line())+
    geom_hline(yintercept = c(1,2, 3), colour = "black")+
    facet_wrap(.~BGC_analysis, ncol = 1, scales = "free_x")
```

### Model Parameters



+ The climate change portfolio was run using predictions up to `r max(timePeriods)+15`.

+ Minimum weight to be included in the portfolio is set at `r minAccept*100`%.

+ The fixed return line is set at `r returnValue`

+ The table below shows the min and max values for each species in the portfolio. Minimum and maximum portfolio ratios can be set for the entire portfolio or by species. These limits will be accounted for in the portfolio often by rebalancing it to include more of an species with similar environmental responses. For example, capping Hw abundance in the portfolio will commonly lead to an increase in the amount of Cw or Ba in the portfolio.

`r ft2`
